---
title: "Hospital Presentation"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")
knitr::opts_chunk$set(echo = TRUE)
```


## 1) The "best" Function

I started by writing a function called best to return the best hospital for a user inputted outcome and state.

```{r}
best <- function(state, outcome){
  options(warn=-1)
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  StateChecker <- unique(data[, "State"])
  if(!(state %in% StateChecker)){stop("Invalid State!")}
  if(outcome == "heart attack"||outcome == "Heart Attack"){outcome = 11}
  if(outcome == "heart failure"||outcome == "Heart Failure"){outcome = 17}
  if(outcome == "pneumonia"||outcome == "Pneumonia"){outcome = 23}
  if(!is.numeric(outcome)){stop("Invalid Outcome!")}
  minpos <- NA
  x <- subset(data, State == state)
  x <- x[order(as.numeric(x[, outcome])), ]
  return (x[1, "Hospital.Name"])
}
```

Here is some example output from the function:

```{r}
best("TX", "heart attack")
best("TX", "heart failure")
best("MD", "heart attack")
best("MD", "pneumonia")
#best("BB", "heart attack")
#best("NY", "hert attack")
```


## 2) The "rankhospital" Function

I then wrote a function that accepts a state, outcome, and rank and returns the hospital in that state with that rank in the corresponding outcome.

```{r}
rankhospital <- function(state, outcome, rank = "best"){
  options(warn=-1)
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  StateChecker <- unique(data[, "State"])
  if(!(state %in% StateChecker)){stop("Invalid State!")}
  if(rank == "best"||rank == "Best"){rank = 1}
  if(outcome == "heart attack"||outcome == "Heart Attack"){outcome = 11}
  if(outcome == "heart failure"||outcome == "Heart Failure"){outcome = 17}
  if(outcome == "pneumonia"||outcome == "Pneumonia"){outcome = 23}
  if(!is.numeric(outcome)){stop("Invalid Outcome!")}
  x <- subset(data, State == state)
  sentVal = FALSE
  x <- x[!is.na(as.numeric(x[, outcome])), ]
  if(rank == "worst"||rank == "Worst"){
    sentVal = TRUE
    rank = 1
  }
  if (rank > nrow(x)){
    return(NA)
  }
  x <- x[order(x[, "Hospital.Name"]), ]
  x <- x[order(as.numeric(x[, outcome]), decreasing = sentVal), ]
  return (x[rank, "Hospital.Name"])
}
```

Here is some example output from this function:

```{r}
rankhospital("TX", "heart failure", 4)
rankhospital("MD", "heart attack", "worst")
rankhospital("MN", "heart attack", 5000)
```


## 3) The "rankall" Function

Finally, I wrote a function accepting an outcome and rank and returning the hospital of that rank in each state with respect to the outcome.

```{r}
rankall <- function(outcome, rank = "best") {
  options(warn=-1)
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  if(rank == "best"||rank == "Best"){rank = 1}
  if(outcome == "heart attack"||outcome == "Heart Attack"){outcome = 11}
  if(outcome == "heart failure"||outcome == "Heart Failure"){outcome = 17}
  if(outcome == "pneumonia"||outcome == "Pneumonia"){outcome = 23}
  if(!is.numeric(outcome)){stop("Invalid Outcome!")}
  States <- unique(data[, "State"])
  States <- States[sort.list(States)]
  output <- c()
  rankIsWorst = FALSE
  if(rank == "worst"||rank == "Worst"||rank == -1){rankIsWorst = TRUE}
  for(state in States){
    x <- subset(data, State == state)
    x <- x[!is.na(as.numeric(x[, outcome])), ]
    x <- x[order(as.numeric(x[, outcome])), ]
    if(rankIsWorst){rank = nrow(x)}
    if(rank > nrow(x)){
      output <- c(output, NA)
    }
    else{
      output <- c(output, x[rank, "Hospital.Name"])
    }  
  }
  return(data.frame(States, output))
}
```

Here is some example output:

```{r}
head(rankall("heart attack", 20), 10)
tail(rankall("pneumonia", "worst"), 3)
tail(rankall("heart failure", 5000), 10)
```